A performance assessment of evolutionary algorithms in a volunteer computing system

### 2016-03-12 PPSN2016/paper.Rnw
Updating the description o the experiments
Which now together with Intro and Abstract take exactly 4 pages as intended. Please @unintendedbear @mariosky @fergunet take a good look at it to see if it's clear enough.

The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
\item A REST (representational state transfer) server that responds to
  HTTP requests made by the browsers. These HTTP requests use JSON for
  encoding information, in this case chromosomes and associates
  information send by the clients and return requests made by the
  server.  The server has the capability to
  run a single experiment, storing the incoming chromosomes in a data structure
  that is reset when the solution is found. This data has a finite
  size that erases the oldest chromosomes once it has filled to
  capacity.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays the fitness and
  other graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client is doing. 
\end{enumerate}

@@

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
Experiment & Time (ms) & \multicolumn{2}{|c|}{IPs} & \multicolumn{2}{|c|}{PUTs} & \#Runs\\
           &           & Total & Contributing & Total & Contributing & \\
\hline
base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 81\\
\hline
graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36\\
graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36\\
\hline
reboot & 513354.5 & 9 & 6 & 545.5 & 62.5 & 40\\
\hline
cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 55\\
\hline
\end{tabular}
%% Initially generated with this code, 
%% <<experiments, cache=FALSE,echo=FALSE>>=
%% experiment.aggregates$Experiment <- c("base","graph1","graph2","graph3","reboot","cachereboot")
%% kable(experiment.aggregates)
%% @ %def

\end{center}
\end{table}
%
Initial experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to \$0.00. Experiments were
announced through a series of posts on Twitter and, in the latest case, Telegram, and
results were published here \cite{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of February. All  %% April --> January ??
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
%Pablo: but it would great at least describe how many times was announced and how long
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 50-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions resets to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that. In fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is indeed negligible. %Pablo: cite

>>>>

The architecture of the {\sf NodIO} system is divided in two tiers:\begin{enumerate}
\item A REST (representational state transfer) server that responds to
  HTTP requests made by the browsers. These HTTP requests use JSON for
  encoding information, in this case chromosomes and associates
  information send by the clients and return requests made by the
  server.  The server has the capability to
  run a single experiment, storing the incoming chromosomes in a data structure
  that is reset when the solution is found. This data has a finite
  size that erases the oldest chromosomes once it has filled to
  capacity.  
\item A client that includes the evolutionary algorithm as
  JavaScript code embedded in a web page that displays the fitness and
  other graphs, some
  additional links, and information on the experiment. This code runs
  an evolutionary algorithm {\em island} that starts with a random
  population, then after every 100 generations, it sends the best individual
  back to the server (via a {\tt PUT} request), and then requests a random
  individual back from the server (via a {\tt GET} request). We have
  kept the number of generations between migrations fixed since it is
  a way of finding out how much real work every client is doing. In
  the version of the server used in this paper, we implemented a
  policy by which the combination IP-fitness value could only be
  inserted once into the cache. That means that if a particular IP
  sends the fitness 99 into the server, it will be accepted only
  once. The main intention of this policy was to avoid {\em
    overlapping} of contributions, with clients still running an {\em
    old} run of the experiment contributing after a new one had
  started. This did not completely avoid these contributions: if the
  clients of the {\em old} run increased their new fitness, it would
  still be allowed. 
\end{enumerate}

@@

For every new individual sent by the clients to the cache, several
pieces of meta data were stored: a time stamp, the client IP, the chromosome and
value, the cache size in that particular moment, and also if that
individual actually updated the cache or not.

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
Experiment & Median Time (ms) & \multicolumn{2}{|c|}{Median IPs} & \multicolumn{2}{|c|}{Median PUTs} & \#Runs\\
           &           & Total & Contributing & Total & Contributing & \\
\hline
base & 120056.0 & 10 & 4 & 256.0 & 21.0 & 81\\
\hline
graph1 & 152619.5 & 9 & 4 & 212.5 & 26.5 & 36\\
graph2 & 1173561.5 & 6 & 4 & 761.5 & 22.0 & 36\\
\hline
reboot & 513354.5 & 9 & 6 & 545.5 & 62.5 & 40\\
\hline
cachereboot & 392917.0 & 12 & 12 & 576.0 & 79.0 & 55\\
\hline
\end{tabular}
%% Initially generated with this code, 
%% <<experiments, cache=FALSE,echo=FALSE>>=
%% experiment.aggregates$Experiment <- c("base","graph1","graph2","graph3","reboot","cachereboot")
%% kable(experiment.aggregates)
%% @ %def

\end{center}
\end{table}
%
Since the point of using volunteer computing system is to cut costs,
all experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to \$0.00. In fact, the NodIO system can be deployed to any
system as long as it can run node.js and access to the filesystem, for
log files, is available. All these logs are available under
request. Processed files are available in the same repository that
hosts this paper, as well as the scripts needed to process them.

Every run had slightly different conditions, although in some cases it
was just a server reboot and a new round of announcements. Each run
batch included at least 30 competions, that is, running until the
solution was found. The number of runs was controlled by polling an
URL that indicated the number of runs made in the present batch so
far. A summary of the experiments and  its results is shown in Table
\ref{tab:runs}. This table shows the median time needed to find the
solution as well as the median number of IPs and PUTs.
